{ \fontfamily{ptm}\selectfont First Order Nonsmooth Optimization: Unifying Nesterov's acceleration and its applications }

### 2025-01-27 Notebook LaTEX/Proposal/presentation.tex
>w<, fixing cringe shit

            Denotes the outerloop algorithm, i.e: Definition \ref{def:lin-catalyst-original} by $\mathbb A$, denote the inner loop algorithm by $\mathbb M$. 

@@

    \subsection{Complexity theoreis with absolute errors}

@@

            The following assumption is crucial and we will return to it. 

>>>>

@@

        \begin{frame}{Notations and Variance Reduced Methods}
            Denotes
                \item the outer loop algorithm, i.e: Definition \ref{def:lin-catalyst-original} by $\mathbb A$, it generates $(x_k, y_k)_{k \ge 1}$; 
                \item the inner loop algorithm by $\mathbb M$ that evaluates $x_k \in \mathcal J_{\kappa^{-1}}^{\epsilon_k} y_{k - 1}$, it generates $(z_{k, t})_{t \ge 0}$. 
            The class of variance reduced methods are often used for $\mathbb M$. 
            Major examples of VRMs include SVRG by Xiao, Zhang \cite{xiao_proximal_2014}, Finito by Defazio et al. \cite{defazio_finito_2014}, SAG by Schmidt et al. \cite{schmidt_minimizing_2017}, and SAGA by \cite{defazio_saga_2014}. 
    \subsection{Complexity theories with absolute errors}

@@

            The following assumption is crucial, and we will return to it. 

### 2025-01-27 Notebook LaTEX/Proposal/proposal.tex
>w<, fixing cringe shit

    \end{align}\label{eqn:example_algorithm}

@@

            $A$ is set to be a positive semi-definite diagonal matrix so the problem admits unique minimizer $x^* = \mathbf 0$ with minimum being zero. 

@@

            The following parameters are used to setup the problem: 

@@

                \item $\epsilon > 0$ is the tolerance value and it's set to be $10^{-10}$. 

@@

\section{Catalyst accelerations and future works}\label{sec:catalyst}
    We start by introducing important assumptions and notations for the Catalyst Meta Acceleration Frameworks. 
    \begin{assumption}\label{ass:catalyst1}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Assume that minimizer exists for $F$ and the minimum is $F^*$. 
        For all $x,y\in \RR^n, \beta > 0$ define the model function: 
        \begin{align*}
            \mathcal M^{\beta^{-1}}_F(x; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{F,\beta^{-1}}(y) := \min_{x\in \RR^n} \mathcal M_F^{\beta^{-1}}(x; y)$. 
        We denote $\mathcal J_{\beta^{-1}F}$ to be the resolvent operator for subgradient of $F$, which is also called the proximal operator. 
    \end{assumption}
    \begin{definition}[Absolute termination criterion C1]\label{def:catalyst-termination-c1}
        Take $F$ as given by Assumption \ref{ass:catalyst1}.
        For $\epsilon > 0, \kappa > 0$ and $x \in \RR^n$, the absolute criterion C1 characterizes the set of inexact proximal iterates by: 
        \begin{align*}
            \mathcal J_{\kappa^{-1} F}^\epsilon (x) := 
            \left\lbrace
                y \in \RR^n \left | \; 
                        \mathcal M^{1/\kappa}_F (y; x) - 
                        \mathcal M^*_{F, 1/\kappa}(x) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
    \end{definition}
    \begin{remark}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}}y = \mathcal J^0_{\beta^{-1}}y$. 
    \end{remark}
    \textbf{The remainder of the proposal is organized as the follow}. 
    Section \ref{ssec:catalyst-intro} gives the main ideas behind the Catalyst Acceleration Framework. 
    It follows Lin et al.'s first paper \cite{lin_universal_2015} on Catalyst Acceleration and summarizes the convergence analysis of the algorithm. 
    Section \ref{ssec:catalyst-second-paper} introduces Lin et al.'s second paper \cite{lin_catalyst_2018} on Catalyst which considers the improved relative error condition stated in Definition \ref{def:catalyst-termination-c2} and discusses its significance. 
    This condition improves the overall complexity and inspired nonconvex gradient based 4WD Catalyst by Paquette \cite{paquette_catalyst_2018}.

@@

        Inspired by accelerated proximal point Method (APPM) from Güler \cite{guler_new_1992}, and Inexact Proximal Point Method (IPPM) of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims   of (APPM) to accelerate Variance Reduced Method (VRM). 

@@

        See Bertsekas's surveys \cite{bertsekas_incremental_2011,bertsekas_incremental_2017} for more context. 

@@

        Next, we introduce the first Catalyst algorithm in \cite{lin_universal_2015}. 

@@

            Let $F:\RR \rightarrow \overline \RR$ be $\mu \ge 0$ strongly convex and closed. 
            Let the initial estimate be $x_0 \in \RR^n$, fix parameters $\kappa > 0$ and $\alpha_0 \in (0, 1]$. 

@@

                Initialize $x_0 = y_0$. Then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ such that: 

@@

        With a fixed regularization parameter $\kappa > 0$, the outer loop (i.e: Definition \ref{def:lin-catalyst-original}) produces $(y_{k}, x_k)_{k \ge k}$ denote it by $\mathbb A$. 

@@

        Lin et al.'s second paper on Catalyst Acceleration \cite{lin_catalyst_2018} describes new ideas to choose the termination criteria for the inner for evaluating $\mathcal J_{\kappa^{-1}F}^{\epsilon_k} y_k$ and gives alternatives to the error sequence $\epsilon_k$ based on those termination criteria. 
        % To elucidate, consider the model function $\mathcal M_F^{\kappa^{-1}}(x; y)$ to be $\mu + \kappa$ strongly convex. 
        % Therefore, it has error bound condition: 
        %     (\forall x\in \RR^n) \quad 
        %     \mathcal M^{1/\kappa}_F(x; y) -
        %     \mathcal M_{F, 1/\kappa}^*(y) 
        %     &\le (\kappa + \mu)\dist
        %             \mathbf 0, \partial \mathcal M^{1/\kappa}_F(x; y)
        %         \right)^2. 
        % By choosing $x$ such that $\dist\left(\mathbf 0, \partial \mathcal M_F^{1/\kappa}(x; y)\right) \le \sqrt{\epsilon}$, it ensures $x \in \mathcal J^\epsilon_{\kappa^{-1}F}(y)$. 
        % Unfortunately, this is a difficult in practice because a full gradient evaluation on the model function $\mathcal M^{1/\kappa}_F(\cdot; y)$ is costly (compare to the small amount required for VRM, which is just the gradient of a few samples.), so Lin suggested alternatives of inner loop termination criteria, and/or upper bounds of inner loop iteration to make Catalyst Acceleration competitive in practice. 
        This section summarizes major results in Lin's second Catalyst paper \cite{lin_catalyst_2018}. 

@@

            Lemma \ref{lemma:sufficient-c1} describes a sufficient conditions to verify the membership of $\bar z \in \mathcal J_{\kappa^{-1}F}^\epsilon (x)$ through the proximal gradient operator on $F:= f_\kappa(z) + g(z)$ which has practical importance because it translates criterion C1 into something implementable, i.e: the proximal gradient operator. 
            For theoretical interests, the absolute and relative criteria C1, C2 places bound on the true error of the gradient of Moreau Envelope at the point $x$. 
            The following are results proved in Lin's second Catalyst paper. 

@@

            \begin{definition}[Catalyst Acceleration with relative inexactness]\label{def:catalyst-relative}

@@

            \begin{theorem}[Outer loop complexity under criterion C2]\label{theorem:catalyst2-outer-loop}
                For the iterates $(x_k)_{k \ge 0}$ generated by algorithm in Definition \ref{def:catalyst-relative}, we have 
                    \item If $\mu > 0$, choose $\alpha_0 = \sqrt{q}$, $\delta_k = \sqrt{q}/(2 - \sqrt{q})$. 
                    Then the iterates $(x_k)_{k \ge 0}$ satisfies $F(x_k) - F^* \le \mathcal O\left(1 - \sqrt{q}/2\right)^k$. 
                    \item If $\mu = 0$, choose $\alpha_0 = 1$, $\delta_k = 1/(k + 1)^2$ satisfies $F(x_k) - F^* \le \mathcal O(k^{-2})$. 

@@

                This is Proposition 8, 9 in Lin et al.'s second Catalyst paper \cite{lin_catalyst_2018}.
                For a precise description of the upper bound, see Theorem 8. 

>>>>

    \end{align}\label{eqn:example-algorithm}

@@

            $A$ is set to be a positive semi-definite diagonal matrix, so the problem admits unique minimizer $x^* = \mathbf 0$ with minimum being zero. 

@@

            The following parameters are used to set up the problem: 

@@

                \item $\epsilon > 0$ is the tolerance value, and it's set to be $10^{-10}$. 

@@

\section{Catalyst accelerations}\label{sec:catalyst}
    We introduce assumptions and notations for the Catalyst Meta Acceleration Frameworks. 
    \begin{assumption}\label{ass:catalyst1}
        Given any $\beta > 0$ and $y \in \RR^n$, and $F: \RR^n\rightarrow \overline \RR$ is $\mu \ge 0$ strongly convex and closed. 
        Assume that minimizer exists for $F$ and the minimum is $F^*$. 
        For all $x,y\in \RR^n, \beta > 0$ define the model function: 
        \begin{align*}
            \mathcal M^{\beta^{-1}}_F(x; y) &:= 
            F(x) + \frac{\beta}{2}\Vert x - y\Vert^2.
        \end{align*}
        We define the Moreau Envelope at $y \in \RR^n$ to be $\mathcal M^*_{F,\beta^{-1}}(y) := \min_{x\in \RR^n} \mathcal M_F^{\beta^{-1}}(x; y)$ and the resolvent operator for subgradient of $F$ to be $\mathcal J_{\beta^{-1}F}$, i.e: which is also called the proximal operator. 
    \end{assumption}
    \begin{definition}[Absolute termination criterion C1]\label{def:catalyst-termination-c1}
        Take $F$ as given by Assumption \ref{ass:catalyst1}.
        For $\epsilon > 0, \kappa > 0$ and $x \in \RR^n$, the absolute criterion C1 characterizes the set of inexact proximal iterates by: 
        \begin{align*}
            \mathcal J_{\kappa^{-1} F}^\epsilon (x) := 
            \left\lbrace
                y \in \RR^n \left | \; 
                        \mathcal M^{1/\kappa}_F (y; x) - 
                        \mathcal M^*_{F, 1/\kappa}(x) \le \epsilon
                \right.
            \right\rbrace. 
        \end{align*}
    \end{definition}
    \begin{remark}
        Setting $\epsilon = 0$, we have the exact definition of the exact resolvent given as $\mathcal J_{\beta^{-1}}y = \mathcal J^0_{\beta^{-1}}y$. 
    \end{remark}
    \textbf{The remainder of this section is organized as the follow}. 
    Section \ref{ssec:catalyst-intro} gives the main ideas behind the Catalyst Acceleration Framework. 
    It follows Lin et al.'s first paper \cite{lin_universal_2015} on Catalyst Acceleration and summarizes the convergence analysis of the algorithm. 
    Section \ref{ssec:catalyst-second-paper} introduces Lin et al.'s second paper \cite{lin_catalyst_2018} on Catalyst which considers the improved relative error condition stated in Definition \ref{def:catalyst-termination-c2}. 
    The relative error condition improves the overall complexity and inspired nonconvex gradient based 4WD Catalyst by Paquette \cite{paquette_catalyst_2018}.

@@

        Inspired by accelerated proximal point Method (APPM) from Güler \cite{guler_new_1992}, and Inexact Proximal Point Method (IPPM) of Rockafellar 1976 \cite{rockafellar_monotone_1976}, Lin et al. \cite{lin_universal_2015} proposed a generic method taking inspirations from the convergence claims of (APPM) to accelerate Variance Reduced Method (VRM). 

@@

        See Bertsekas's surveys \cite{bertsekas_incremental_2011,Bertsekas_incremental_2017} for more context. 

@@

        The following definition introduces the first Catalyst algorithm in \cite{lin_universal_2015}. 

@@

            Let $F:\RR \rightarrow \overline \RR$ be $\mu \ge 0$ strongly convex and closed, $x_0 \in \RR^n$ be the initial condition and $\kappa > 0$, $\alpha_0 \in (0, 1]$ are fixed parameters for the algorithm. 

@@

                Initialize $x_0 = y_0$. Then the algorithm generates $(x_k, y_k)_{k\ge 0}$ for all $k \ge 1$ by the procedures: 

@@

        With a fixed regularization parameter $\kappa > 0$, the outer loop (i.e: Definition \ref{def:lin-catalyst-original}) produces $(y_{k}, x_k)_{k \ge 1}$ denote it by $\mathbb A$. 

@@

        Lin et al.'s second paper on Catalyst Acceleration \cite{lin_catalyst_2018} describes the new option of relative termination criteria (which we stated here in Definition \ref{def:catalyst-termination-c2}) for $\mathbb M$ to evaluate the inexact solutions of $\mathcal J_{\kappa^{-1}F}^{\epsilon_k} y_k$. 
        This condition allows the the use of relative error sequence $\delta_k$ which has advantages that: it gives easier convergence anlaysis, it doesn't need knowledge of $F^*$ and it can be generalize to the nonconvex case. 
        This section summarizes major results in Lin et al's second Catalyst paper \cite{lin_catalyst_2018}. 

@@

            Lemma \ref{lemma:sufficient-c1} describes a sufficient conditions to verify the membership of $\bar z \in \mathcal J_{\kappa^{-1}F}^\epsilon (x)$ through the proximal gradient operator on $F:= f_\kappa(z) + g(z)$ which has practical importance because it translate criterion C1 into a implementable criteria in the algorithm with the proximal gradient operator. 
            It contribute to theoretical analysis because it bounds on the true error on the gradient of Moreau Envelope at the point $x$. 
            The following results are proved in Lin et al second Catalyst paper \cite{lin_catalyst_2018}. 

@@

            \begin{definition}[Catalyst Acceleration all in one]\label{def:catalyst-all-in-one}

@@

            \begin{assumption}[Warm start conditions for relative error C2]
            \;\label{prop:warm-start-with-c2}\\ 
            Assume

            \begin{theorem}[Outer loop complexity under criterion C2]\; \label{theorem:catalyst2-outer-loop}\\
                For the iterates $(x_k)_{k \ge 0}$ generated by algorithm in Definition \ref{def:catalyst-all-in-one}, we have 
                \begin{enumerate}
                    \item If $\mu > 0$, choose $\alpha_0 = \sqrt{q}$, $\delta_k = \sqrt{q}/(2 - \sqrt{q})$. 
                    Then the iterates $(x_k)_{k \ge 0}$ satisfies $F(x_k) - F^* \le \mathcal O\left(1 - \sqrt{q}/2\right)^k$. 
                    \item If $\mu = 0$, choose $\alpha_0 = 1$, $\delta_k = 1/(k + 1)^2$ satisfies $F(x_k) - F^* \le \mathcal O(k^{-2})$. 
                \end{enumerate}
            \end{theorem}
            \begin{remark}
                This is Proposition 8, 9 in Lin et al.'s second Catalyst paper \cite{lin_catalyst_2018}.
                For a precise description of the upper bound, see Theorem 8. 
            \end{remark}
            \begin{theorem}{Inner loop complexity with relative error C2}

@@
